---
title: "Motor Trend"
author: "Xiaowen Wang"
date: "May 24, 2015"
output: pdf_document
---
# Summary 
In general, it's not significant to say manual or automatic car is better. Include more features, we can tell for lower feathers, manual is better for mpg, while for higher features, automatic is better for mpg. 
```{r,echo=FALSE}
data(mtcars)
```  
# Automatic versus Manual
Devide the data into two groups, g0 is automatic transmission and g1 is manual transmission.
```{r,echo=FALSE}
g0 <- subset(mtcars,am==0)
g1 <- subset(mtcars,am==1)
```  
First, we want to plot mpg of two type. 
```{r,echo=FALSE}
par(mfrow=c(1,1),mar=c(6,6,6,6))
plot(y=mtcars$mpg,x=mtcars$am,main="mpg of automatic and manual",xlab="am",ylab="mpg",col="green",pch=17)
```  
It seems that manual is better for mpg. Then let's do the hypothesis to see if it's true. Our null hypothesis is that it makes no difference whether it's manual or autumatic.
Ho: mu0 = mu1
Ha: mu0 < mu1
```{r,echo=FALSE}
t.test(g0,g1,paired=FALSE,var.equal=FALSE)
```
AS p > 0.05 and t value is small, we're not able to reject the null hypothesis. That's say, you can't say manual is better for mpg, compared to automatic. The difference might be caused by perticular driving hobbit or other feathers of the car.  

# Regression Model
We notice there seem to be regression models outcome as mpg, income as disp,hp wt or carb. Let's first do linear regression one by one and then do a maltivarable regression.   

## Linear Regression
```{r,echo=FALSE}
c00 <- cor(x=g0$disp,y=g0$mpg)
l00 <- lm(mpg~disp,g0)
c01 <- cor(x=g1$disp,y=g1$mpg)
l01 <- lm(mpg~disp,g1)
c10 <-cor(x=g0$hp,y=g0$mpg)
l10 <- lm(mpg~hp,g0)
c11 <- cor(x=g1$hp,y=g1$mpg)
l11 <- lm(mpg~hp,g1)
c20 <- cor(x=g0$wt,y=g0$mpg)
l20 <- lm(mpg~wt,g0)
c21 <- cor(x=g1$wt,y=g1$mpg)
l21 <- lm(mpg~wt,g1)
c30 <- cor(x=g0$carb,y=g0$mpg)
l30 <- lm(mpg~carb,g0)
c31 <- cor(x=g1$carb,y=g1$mpg)
l31 <- lm(mpg~carb,g1)
m <- matrix(data=c(c00,c01,c10,c11,c20,c21,c30,c31),nrow=4,ncol=2,byrow=TRUE)
names(m) <- c("auto","manual")
m
``` 

## Multivariable Regression
```{r}
l0 <- lm(mpg~disp+hp+wt+carb,g0);l0
l1 <- lm(mpg~disp+hp+wt+carb,g1);l1
```
According to the above regression model, it's convincible that disp,hp, wt and carb has a linear relationship with mpg. According to the multivariable regression, we can see automatic has a lower starting point of mpg, but features has less affects on the mpg compared to manual car. For lower disp,hp,wt and carb, manual cars are better for mpg; For higher disp,hp, wt and carb, automatic cars are better for mpg.  

# Appendix 
````{r,echo=FALSE}
par(mfrow=c(3,3),mar=c(2,2,2,2))
plot(y=g0$mpg,x=g0$cyl,main="cyl");plot(y=g0$mpg,x=g0$disp,main="disp");plot(y=g0$mpg,x=g0$hp,main="hp");plot(y=g0$mpg,x=g0$drat,main="drat");plot(y=g0$mpg,x=g0$wt,main="wt");plot(y=g0$mpg,x=g0$qsec,main="qsec");plot(y=g0$mpg,x=g0$vs,main="vs");plot(y=g0$mpg,x=g0$gear,main="gear");plot(y=g0$mpg,x=g0$carb,main="carb")
mtext("Features for Automatic Cars",outer=TRUE,cex=1.5)

par(mfrow=c(3,3))
plot(y=g1$mpg,x=g1$cyl,main="cyl");plot(y=g1$mpg,x=g1$disp,main="disp");plot(y=g1$mpg,x=g1$hp,main="hp");plot(y=g1$mpg,x=g1$drat,main="drat");plot(y=g1$mpg,x=g1$wt,main="wt");plot(y=g1$mpg,x=g1$qsec,main="qsec");plot(y=g1$mpg,x=g1$vs,main="vs");plot(y=g1$mpg,x=g1$gear,main="gear");plot(y=g1$mpg,x=g1$carb,main="carb")
mtext("Features for Manual Cars", outer = TRUE, cex = 1.5)

par(mfrow=c(3,3))
plot(y=mtcars$mpg,x=mtcars$cyl,main="cyl");plot(y=mtcars$mpg,x=mtcars$disp,main="disp");plot(y=mtcars$mpg,x=mtcars$hp,main="hp");plot(y=mtcars$mpg,x=mtcars$drat,main="drat");plot(y=mtcars$mpg,x=mtcars$wt,main="wt");plot(y=mtcars$mpg,x=mtcars$qsec,main="qsec");plot(y=mtcars$mpg,x=mtcars$vs,main="vs");plot(y=mtcars$mpg,x=mtcars$gear,main="gear");plot(y=mtcars$mpg,x=mtcars$carb,main="carb")
mtext("Features for Both", outer = TRUE, cex = 1.5)

par(mfrow=c(2,2))
plot(l0)
plot(l1)
```
# 